---
title: "tspresent"
author: "Zhejia Dong"
date: "2023-04-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,eval = TRUE,cache=TRUE,warning = FALSE,message = FALSE)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(fig.height = 4)
knitr::opts_chunk$set(fig.width = 6)
knitr::opts_chunk$set(fig.align="center")
```

```{r libarary, include = FALSE, echo =TRUE}
library(tidyr) #Allows for us to manipulate the data structure
library(data.table) #Allows for us to manipulate the data structure
library(ggplot2) #this makes better looking plots in R
library(dplyr)
library(haven)
library(astsa)
library(TTR)
library(forecast)
library(broom.mixed)
library(cowplot)
library(knitr)
library(TSA)
library(energy)
library(plotrix)
library(kableExtra)
theme_set(theme_minimal())
```


# MA(2) - second order moving average process

$$
Y_t = e_t + \theta_1 e_{t-1} + \theta_2 e_{t-2}
$$
where $e_{i}$ is white noise from $N(0, \sigma_e^2)$. Then we can get the variance and covariance as follows:

$$
\begin{align*}
\gamma_0 & = Var(\gamma_t) = Var(e_t + \theta_1 e_{t-1} + \theta_2 e_{t-2}) = (1+\theta^2_1 + \theta^2_2) \sigma^2_e\\
\gamma_1 & = Cov(Y_t, Y_{t-1}) \\
& = Cov(e_t + \theta_1 e_{t-1} + \theta_2 e_{t-2}, e_{t-1} + \theta_1 e_{t-2} + \theta_2 e_{t-3} )\\
& = (\theta_1 + \theta_1\theta_2)\sigma^2_e\\
\gamma_2& = Cov(Y_t, Y_{t-2})\\
& =Cov(e_t + \theta_1 e_{t-1} + \theta_2 e_{t-2}, e_{t-2} + \theta_1 e_{t-3} + \theta_2 e_{t-4} )\\
& = \theta_2\sigma^2_e
\end{align*}
$$

For a MA(2) time series {$Y_t$},$t = 1,2,\cdots, n$, the variance matrix should be

\[
\begin{bmatrix}
    1 + \theta_1^2 + \theta_2^2 & \theta_1 + \theta_1\theta_2 &  \theta_2 & 0 & \cdots & 0 \\
    \theta_1 + \theta_1\theta_2 & 1 + \theta_1^2 + \theta_2^2 &  \theta_1 + \theta_1\theta_2 & \theta_2 & \cdots & 0 \\
     \theta_2 &  \theta_1 + \theta_1\theta_2 & 1 + \theta_1^2 + \theta_2^2 & \theta_1 + \theta_1\theta_2 & \cdots & 0 \\
    0 &  \theta_2 &  \theta_1 + \theta_1\theta_2 & 1 + \theta_1^2 + \theta_2^2 & \cdots & 0 \\
    \vdots & \vdots & \vdots & \vdots & \ddots & \vdots \\
    0 & 0 & 0 & 0 & \cdots & 1 + \theta_1^2 + \theta_2^2
\end{bmatrix} \sigma^2_e
\]


We then start to generate the MA(2). The parameter set to be $(\theta_1, \theta_2) = (0.7, 0.9)$. white noise {$e_t$} from standard normal distribution with zero mean and unit variance. The sample size is set to 1000.



```{r}
set.seed(98234) 
theta1 <- 0.7
theta2 <- 0.3
popn <- 500
eps <- rnorm(popn+2, mean = 0, sd = 1)
ts <- c()

# Creating example series
for(i in 1:popn){
  ts[i] <- eps[i+2] + theta1 * eps[i+1]  + theta2 * eps[i]
}
ts.plot(ts)
### or use arima.sim 
## arima.sim(  n = 500, list(ma = c(0.7,0.3)), sd = 1)


```



```{r, eval=FALSE}


popn <- 1000
n.sim <- 500
ma.per <- c(0.9,0.9)
ma.process <- function(ma.per, popn, mu = 0, sd = 1){
  q <- length(ma.per)
  eps.x <- rnorm((popn + q), mean = mu, sd = sd)
  eps.y <- rnorm((popn + q), mean = mu, sd = sd)
  x <- c()
  y <- c()
  for(i in (1+q):(popn+q)){
    x[i] <- sum(eps.x[i:(i-q)] * c(1, ma.per[1:q]))
    y[i] <- sum(eps.y[i:(i-q)] * c(1, ma.per[1:q]))
  }
  x <- x[(1+q) : (popn + q)]
  y <- y[(1+q) : (popn + q)]
  
  return(list("x" = x, "y" = y, "ma.par" = ma.per,
              "mu" = mu, "sd" = sd))
}

x.ts <- matrix(0, nrow = n.sim, ncol = popn) # per row is one replication
y.ts <- matrix(0, nrow = n.sim, ncol = popn)


for(i in 1: n.sim){
  per.process <-ma.process(ma.per, popn)
  x.ts[i,] <- per.process$x
  y.ts[i,] <- per.process$y
}



```

```{r result, eval=FALSE}
cor.coef <- c()
dis.core <- c()
dis.p <- c()
lm.p <- c()
lm.beta <- c()
z.beta <- c()


for(i in 1:n.sim){
  cor.coef[i] <- cor(x.ts[i,], y.ts[i,])
  dis.core[i] <- dcor(x.ts[i,], y.ts[i,])
  dd <- dcorT.test(x.ts[i,], y.ts[i,])
  dis.p[i] <-dd$p.value
  lt <- lm(y.ts[i,] ~ x.ts[i,])
  ls <- summary(lt)
  lm.beta[i] <- ls$coefficients[2,"Estimate"]
  lm.p[i] <- ls$coefficients[2,4] ## p value for linear regression test, p value.
  z.beta[i] <- lm.beta[i]/ (ls$coefficients[2,"Std. Error"] * sd(x.ts[i,]))
}



```

## First simulation MA(2) (0.7,0.3)

```{r, echo=FALSE}
rm(list = ls())
load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/first.sim.RData")
#load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/second.sim.RData")
```

- cross correlation function plot
```{r}

ccf(y.ts[1,], x.ts[1,])

```

- Correlation coefficients (ccf at lag 0)
- Distance correlation
- Standard beta correlation
- Rejection rates of linear and distance correlaion test

```{r}
load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/first.sim.RData")
hist(cor.coef, main = "Correlation coefficients for MA(2)",
     xlab = "Correlation coefficient", col = "lightblue",prob=T)

hist(dis.core, main = "Distance correlations for MA(2)",
     xlab = "Distance correlation", col = "lightblue", prob=T)

hist(z.beta, main = expression(bold(paste("Standardized ", beta, " coefficients for MA(2)"))),
     xlab = expression(paste("Standardized ", beta, " coefficients")) ,
     col = "lightblue",prob=T)


data.frame( "linear test" = mean(lm.p < 0.05),
            "dis test" = mean(dis.p < 0.05)) %>% 
  kbl( booktabs=T, escape=T, align = "c",digits = 3,
                caption = "Rejection rates for linear and distance correlation test") %>%
  kable_styling(full_width =F, latex_options = c("HOLD_position")) %>% 
  kable_material(c("striped")) 
```

## Second simulation MA(2) (0.9,0.9) highly correlated

```{r, echo=FALSE}
rm(list = ls())
#load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/first.sim.RData")
load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/second.sim.RData")
```

- cross correlation function plot

```{r}
ccf(y.ts[1,], x.ts[1,])

```

- Correlation coefficients (ccf at lag 0)
- Distance correlation
- Standard beta correlation
- Rejection rates of linear and distance correlation test


```{r, echo=FALSE}
load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/second.sim.RData")
hist(cor.coef, main = "Correlation coefficients for MA(2)",
     xlab = "Correlation coefficient", col = "lightblue",prob=T)

hist(dis.core, main = "Distance correlations for MA(2)",
     xlab = "Distance correlation", col = "lightblue", prob=T)

hist(z.beta, main = expression(bold(paste("Standardized ", beta, " coefficients for MA(2)"))),
     xlab = expression(paste("Standardized ", beta, " coefficients")) ,
     col = "lightblue",prob=T)


data.frame( "linear test" = mean(lm.p < 0.05),
            "dis test" = mean(dis.p < 0.05)) %>% 
  kbl( booktabs=T, escape=T, align = "c",digits = 3,
                caption = "Rejection rates for linear and distance correlation test") %>%
  kable_styling(full_width =F, latex_options = c("HOLD_position")) %>% 
  kable_material(c("striped")) 
```



## AR(2)

Then we consider auto regressive process.

$$Y_t = \phi_1 Y_{t-1} + \phi_2 Y_{t-2} + e_t $$
AR characteristic polynomial $\phi(x) = 1 -\phi_1x -\phi_2 x^2$

A simple sample：

```{r}
set.seed(123456)
popn <- 1000
ar.par <- c(0.7, 0.3)
p <- length(ar.par)
eps <- rnorm( popn + p, 0, 1)

ts <- rnorm(popn+p, 0, 1)

for(i in (p+1) : (popn+p)){
  ts[i] <-   sum(ar.par * ts[(i-1):(i-p)]) + eps[i] 
}

ts.plot(ts[(p+1):(popn + p)])
```

## Third simulation AR(2) (0.7,0.3)

```{r, echo=FALSE,eval=FALSE}
popn <- 1000
n.sim <- 500
ar.par <- c(0.7, 0.3)
p <- length(ar.par)


ar.process <- function(ar.par, popn, mu = 0, sd = 1){
  p <- length(ar.par)
  eps.x <- rnorm((popn + p), mean = mu, sd = sd)
  eps.y <- rnorm((popn + p), mean = mu, sd = sd)
  x <- rnorm(popn+p, 0, 1)
  y <- rnorm(popn+p, 0, 1)
  for(i in (1+p):(popn+p)){
    x[i] <-  sum(ar.par * x[(i-1):(i-p)]) + eps.x[i]  
    y[i] <-  sum(ar.par * x[(i-1):(i-p)])  + eps.y[i]
  }
  x <- x[(1+p) : (popn + p)]
  y <- y[(1+p) : (popn + p)]
  
  return(list("x" = x, "y" = y, "ar.par" = ar.par,
              "mu" = mu, "sd" = sd))
}

x.ts <- matrix(0, nrow = n.sim, ncol = popn) # per row is one replication
y.ts <- matrix(0, nrow = n.sim, ncol = popn)

for(i in 1: n.sim){
  per.process <- ar.process(ar.par, popn)
  x.ts[i,] <- per.process$x
  y.ts[i,] <- per.process$y
}


cor.coef <- c()
dis.core <- c()
dis.p <- c()
lm.p <- c()
lm.beta <- c()
z.beta <- c()


for(i in 1:n.sim){
  cor.coef[i] <- cor(x.ts[i,], y.ts[i,])
  dis.core[i] <- dcor(x.ts[i,], y.ts[i,])
  dd <- dcorT.test(x.ts[i,], y.ts[i,])
  dis.p[i] <-dd$p.value
  lt <- lm(y.ts[i,] ~ x.ts[i,])
  ls <- summary(lt)
  lm.beta[i] <- ls$coefficients[2,"Estimate"]
  lm.p[i] <- ls$coefficients[2,4] ## p value for linear regression test, p value.
  z.beta[i] <- lm.beta[i]/ (ls$coefficients[2,"Std. Error"] * sd(x.ts[i,]))
}



```

```{r, echo=FALSE}
rm(list = ls())
#load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/first.sim.RData")
#load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/second.sim.RData")
load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/third.sim.RData")
```

- cross correlation function plot

```{r echo=FALSE}
ccf(y.ts[1,], x.ts[1,])

```

- Correlation coefficients (ccf at lag 0)
- Distance correlation
- Standard beta correlation
- Rejection rates of linear and distance correlation test


```{r echo=FALSE}
load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/third.sim.RData")
hist(cor.coef, main = "Correlation coefficients for AR(2)",
     xlab = "Correlation coefficient", col = "lightblue",prob=T)

hist(dis.core, main = "Distance correlations for AR(2)",
     xlab = "Distance correlation", col = "lightblue", prob=T)

hist(z.beta, main = expression(bold(paste("Standardized ", beta, " coefficients for AR(2)"))),
     xlab = expression(paste("Standardized ", beta, " coefficients")) ,
     col = "lightblue",prob=T)


data.frame( "linear test" = mean(lm.p < 0.05),
            "dis test" = mean(dis.p < 0.05)) %>% 
  kbl( booktabs=T, escape=T, align = "c",digits = 3,
                caption = "Rejection rates for linear and distance correlation test") %>%
  kable_styling(full_width =F, latex_options = c("HOLD_position")) %>% 
  kable_material(c("striped")) 
```

## Forth simulation AR(2) (0.9,-0.9)
```{r, echo=FALSE}
rm(list = ls())
#load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/first.sim.RData")
#load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/second.sim.RData")
#load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/third.sim.RData")
load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/forth.sim.RData")
```

- cross correlation function plot

```{r echo=FALSE}
ccf(y.ts[1,], x.ts[1,])

```

- Correlation coefficients (ccf at lag 0)
- Distance correlation
- Standard beta correlation
- Rejection rates of linear and distance correlation test


```{r echo=FALSE}
load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/forth.sim.RData")
hist(cor.coef, main = "Correlation coefficients for AR(2)",
     xlab = "Correlation coefficient", col = "lightblue",prob=T)


hist(dis.core, main = "Distance correlations for AR(2)",
     xlab = "Distance correlation", col = "lightblue", prob=T)

hist(z.beta, main = expression(bold(paste("Standardized ", beta, " coefficients for AR(2)"))),
     xlab = expression(paste("Standardized ", beta, " coefficients")) ,
     col = "lightblue",prob=T)


data.frame( "linear test" = mean(lm.p < 0.05),
            "dis test" = mean(dis.p < 0.05)) %>% 
  kbl( booktabs=T, escape=T, align = "c",digits = 3,
                caption = "Rejection rates for linear and distance correlation test") %>%
  kable_styling(full_width =F, latex_options = c("HOLD_position")) %>% 
  kable_material(c("striped")) 
```


## More simulation for AR(0.3,0.3)

```{r, echo=FALSE}
rm(list = ls())
#load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/first.sim.RData")
#load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/second.sim.RData")
#load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/third.sim.RData")
load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/fifth.sim.RData")

ccf(y.ts[1,], x.ts[1,])

hist(cor.coef, main = "Correlation coefficients for AR(2)",
     xlab = "Correlation coefficient", col = "lightblue",prob=T)

hist(dis.core, main = "Distance correlations for AR(2)",
     xlab = "Distance correlation", col = "lightblue", prob=T)

hist(z.beta, main = expression(bold(paste("Standardized ", beta, " coefficients for AR(2)"))),
     xlab = expression(paste("Standardized ", beta, " coefficients")) ,
     col = "lightblue",prob=T)


data.frame( "linear test" = mean(lm.p < 0.05),
            "dis test" = mean(dis.p < 0.05)) %>% 
  kbl( booktabs=T, escape=T, align = "c",digits = 3,
                caption = "Rejection rates for linear and distance correlation test") %>%
  kable_styling(full_width =F, latex_options = c("HOLD_position")) %>% 
  kable_material(c("striped")) 
```




## More simulation for AR(-0.3, -0.3)

```{r, echo=FALSE}
rm(list = ls())

load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/six.sim(-0.3,-0.3).RData")

ccf(y.ts[1,], x.ts[1,])

hist(cor.coef, main = "Correlation coefficients for AR(2)",
     xlab = "Correlation coefficient", col = "lightblue",prob=T)

hist(dis.core, main = "Distance correlations for AR(2)",
     xlab = "Distance correlation", col = "lightblue", prob=T)

hist(z.beta, main = expression(bold(paste("Standardized ", beta, " coefficients for AR(2)"))),
     xlab = expression(paste("Standardized ", beta, " coefficients")) ,
     col = "lightblue",prob=T)


data.frame( "linear test" = mean(lm.p < 0.05),
            "dis test" = mean(dis.p < 0.05)) %>% 
  kbl( booktabs=T, escape=T, align = "c",digits = 3,
                caption = "Rejection rates for linear and distance correlation test") %>%
  kable_styling(full_width =F, latex_options = c("HOLD_position")) %>% 
  kable_material(c("striped")) 
```



## More Simulation for AR(0.5, 0.5)

```{r, echo=FALSE}
rm(list = ls())

load("~/Dropbox (Brown)/2023 Spring RAship/Spring23_RA/ts present/data/seventh.sim(0.5,0.5).RData")

ccf(y.ts[1,], x.ts[1,])

hist(cor.coef, main = "Correlation coefficients for AR(2)",
     xlab = "Correlation coefficient", col = "lightblue",prob=T)

hist(dis.core, main = "Distance correlations for AR(2)",
     xlab = "Distance correlation", col = "lightblue", prob=T)

hist(z.beta, main = expression(bold(paste("Standardized ", beta, " coefficients for AR(2)"))),
     xlab = expression(paste("Standardized ", beta, " coefficients")) ,
     col = "lightblue",prob=T)


data.frame( "linear test" = mean(lm.p < 0.05),
            "dis test" = mean(dis.p < 0.05)) %>% 
  kbl( booktabs=T, escape=T, align = "c",digits = 3,
                caption = "Rejection rates for linear and distance correlation test") %>%
  kable_styling(full_width =F, latex_options = c("HOLD_position")) %>% 
  kable_material(c("striped")) 
```



```{r eval=FALSE}
set.seed(923)
popn <- 1000
n.sim <- 500
ar.par <- c(0.5, 0.5)
p <- length(ar.par)


ar.process <- function(ar.par, popn, mu = 0, sd = 1){
  p <- length(ar.par)
  eps.x <- rnorm((popn + p), mean = mu, sd = sd)
  eps.y <- rnorm((popn + p), mean = mu, sd = sd)
  x <- rnorm(popn+p, 0, 1)
  y <- rnorm(popn+p, 0, 1)
  for(i in (1+p):(popn+p)){
    x[i] <-  sum(ar.par * x[(i-1):(i-p)]) + eps.x[i]  
    y[i] <-  sum(ar.par * x[(i-1):(i-p)])  + eps.y[i]
  }
  x <- x[(1+p) : (popn + p)]
  y <- y[(1+p) : (popn + p)]
  
  return(list("x" = x, "y" = y, "ar.par" = ar.par,
              "mu" = mu, "sd" = sd))
}

x.ts <- matrix(0, nrow = n.sim, ncol = popn) # per row is one replication
y.ts <- matrix(0, nrow = n.sim, ncol = popn)

for(i in 1: n.sim){
  per.process <- ar.process(ar.par, popn)
  x.ts[i,] <- per.process$x
  y.ts[i,] <- per.process$y
}


cor.coef <- c()
dis.core <- c()
dis.p <- c()
lm.p <- c()
lm.beta <- c()
z.beta <- c()


for(i in 1:n.sim){
  cor.coef[i] <- cor(x.ts[i,], y.ts[i,])
  dis.core[i] <- dcor(x.ts[i,], y.ts[i,])
  dd <- dcorT.test(x.ts[i,], y.ts[i,])
  dis.p[i] <-dd$p.value
  lt <- lm(y.ts[i,] ~ x.ts[i,])
  ls <- summary(lt)
  lm.beta[i] <- ls$coefficients[2,"Estimate"]
  lm.p[i] <- ls$coefficients[2,4] ## p value for linear regression test, p value.
  z.beta[i] <- lm.beta[i]/ (ls$coefficients[2,"Std. Error"] * sd(x.ts[i,]))
}


```


